### 2025-12-17 extra/paper-cardinalR-code.Rmd
Added alt-txt for cone data projections

---
title: "Sample cardinalR"
author: "Chat-GPT"
date: "2025-09-10"
format: pdf
---

### Cone

To simulate a cone-shaped structure in arbitrary dimensions (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:cone)"} else { "Figure \\@ref(fig:cone-proj)"})`), we define a function `gen_cone(n, p, h, ratio)`, which creates a high-dimensional cone with options for a sharp or blunted apex, allowing for a dense concentration of points near the tip.

This function generates $n$ points in $p\text{-}D$, where the last dimension, $X_p$, represents the height along the cone's axis, and the first $p-1$ dimensions define a shrinking hyperspherical cross-section toward the tip. Heights are sampled from a truncated exponential distribution, $X_p \sim \text{Exp}(\lambda = 2/h)$, truncated to the interval $[0, h]$, producing a higher density of points near the tip. At each height $X_p$, the radius of the cross-section increases linearly from base to tip according to $r = r_{\text{min}} + (r_{\text{max}} - r_{\text{min}}) X_p / h$, where $r_{\text{min}} = \text{ratio} \in [0, 1]$ and $r_{\text{max}} = 1$.

For each point, a direction in the first $p-1$ dimensions is sampled uniformly on a $(p-1)$-dimensional hypersphere using generalized spherical coordinates. The radial coordinates are scaled by the height-dependent radius $r$, producing the conical taper. In three dimensions ($p = 3$), this results in a classical $3\text{-}D$ cone, while for $p > 3$, additional dimensions provide a smooth embedding into higher-dimensional space, preserving the conical structure.

Cone-shaped structures appear in particle dispersions, light beams, and tapering processes, where spread decreases along one axis. They are also used to benchmark clustering and dimensionality reduction methods [@hadsell2006].

```{r cone-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from the $3\\text{-}D$ \\texttt{cone} data. Points are concentrated near the tip along the height dimension, while the radius of the hyperspherical cross-section decreases linearly toward the apex. These projections show how the conical geometry is preserved.", fig.pos="H", fig.pos="!ht", fig.width=12, fig.height=4}
cone <- gen_cone(n = 1000, p = 4, h = 5, ratio = 0.5)

# Add plot label

interior_annotation <- function(label, position = c(0.92, 0.92), cex = 1, col="grey70") {
  annotation_custom(grid::textGrob(label = label,
                                   x = unit(position[1], "npc"), y = unit(position[2], "npc"),
                                   gp = grid::gpar(cex = cex, col=col)))
}

# Scale high-d data

# Center the data by subtracting the mean of each column
center_data <- function(data) {
  center_values <- colMeans(data)
  data_centered <- sweep(data, 2, center_values, FUN = "-")  # subtract means
  data_centered
}

# Function to scale data manually
# scale_data_manual <- function(data) {
#   # Step 1: Center the data (mean 0)
#   data_centered <- center_data(data)
#
#   # Step 2: Calculate the standard deviation of each dimension
#   sds <- apply(data_centered, 2, sd)
#
#   # Step 3: Scale each dimension to have the range [0, 1]
#   data_scaled <- apply(data_centered, 2, function(col) col / max(abs(col)))
#
#   # Step 4: Scale dimensions according to their variation
#   # The dimension with the highest standard deviation is scaled to [-1, 1]
#   # Other dimensions are scaled to smaller ranges based on their standard deviations
#   max_sd <- max(sds)
#
#   # Normalize the standard deviations to get scaling factors
#   scaling_factors <- sds / max_sd
#
#   for (i in seq_along(scaling_factors)) {
#     data_scaled[, i] <- data_scaled[, i] * scaling_factors[i]
#   }
#
#   # Combine the scaled data with the 'type' column and return as a tibble
#   data_scaled <- as_tibble(data_scaled)
#
#   return(data_scaled)
# }

# Get projection

get_projection <- function(projection, centered_data, axis_param) {

  projected <- as.matrix(centered_data) %*% projection
  projected_df <- projected |>
    tibble::as_tibble(.name_repair = "unique") |>
    dplyr::rename(c("proj1" = "...1",
                    "proj2" = "...2")) |>
    dplyr::mutate(ID = dplyr::row_number())

  limits <- axis_param$limits
  axis_scaled <- axis_param$axis_scaled
  axis_pos_x <- axis_param$axis_pos_x
  axis_pos_y <- axis_param$axis_pos_y
  threshold <- axis_param$threshold

  axes_obj <- gen_axes(
    proj = projection * axis_scaled,
    limits = limits,
    axis_pos_x = axis_pos_x,
    axis_pos_y = axis_pos_y,
    axis_labels = names(centered_data),
    threshold = threshold)

  axes <- axes_obj$axes
  circle <- axes_obj$circle

  return(list(projected_df = projected_df,
              axes = axes,
              circle = circle))

}

# Plot projection
plot_proj <- function(proj_obj,
                      point_param = c(1.5, 0.5, "#000000"), # size, alpha, color
                      plot_limits, title, cex = 2,
                      position = c(0.92, 0.92),
                      axis_text_size = 3,
                      is_color = FALSE) {

  projected_df <- proj_obj$projected_df
  axes <- proj_obj$axes
  circle <- proj_obj$circle

  if(is_color == FALSE) {

    initial_plot <- ggplot() +
      geom_point(
        data = projected_df,
        aes(
          x = proj1,
          y = proj2),
        size = as.numeric(point_param[1]),
        alpha = as.numeric(point_param[2]),
        color = point_param[3])

  } else {

    projected_df <- projected_df |>
      dplyr::mutate(cluster = proj_obj$cluster)

    initial_plot <- ggplot() +
      geom_point(
        data = projected_df,
        aes(
          x = proj1,
          y = proj2,
          color = cluster),
        size = as.numeric(point_param[1]),
        alpha = as.numeric(point_param[2]))

  }

  initial_plot <- initial_plot +
    geom_segment(
      data=axes,
      aes(x=x1, y=y1, xend=x2, yend=y2),
      colour="grey70") +
    geom_text(
      data=axes,
      aes(x=x2, y=y2),
      label=rownames(axes),
      colour="grey50",
      size = axis_text_size) +
    geom_path(
      data=circle,
      aes(x=c1, y=c2), colour="grey70") +
    xlim(plot_limits) +
    ylim(plot_limits) +
    interior_annotation(title, position, cex = cex)

  initial_plot

}

# Generate axes

gen_axes <- function(proj, limits = 1, axis_pos_x = NULL, axis_pos_y = NULL, axis_labels, threshold) {

  axis_scale <- limits/6

  if (is.null(axis_pos_x)) {

    axis_pos_x <- -2/3 * limits

  }

  if (is.null(axis_pos_y)) {

    axis_pos_y <- -2/3 * limits

  }

  adj <- function(x, axis_pos) axis_pos + x * axis_scale
  axes <- data.frame(x1 = adj(0, axis_pos_x),
                     y1 = adj(0, axis_pos_y),
                     x2 = adj(proj[, 1], axis_pos_x),
                     y2 = adj(proj[, 2], axis_pos_y))

  rownames(axes) <- axis_labels

  ## To remove axes
  axes <- axes |>
    mutate(distance = sqrt((x2 - x1)^2 + (y2 - y1)^2)) |>
    filter(distance >= threshold)

  theta <- seq(0, 2 * pi, length = 50)
  circle <- data.frame(c1 = adj(cos(theta), axis_pos_x),
                       c2 = adj(sin(theta), axis_pos_y))

  return(list(axes = axes, circle = circle))

}

plot_hbe <- function(error_df) {

  ggplot(error_df,
         aes(x = a1,
             y = HBE,
             colour = method)) +
    geom_point(size = 0.8) +
    geom_line(linewidth = 0.3) +
    ylab("HBE") +
    xlab(expression(paste("binwidth (", a[1], ")"))) +
    theme_minimal() +
    theme(panel.border = element_rect(fill = 'transparent'),
          plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
          axis.ticks.x = element_line(),
          axis.ticks.y = element_line(),
          legend.position = "none",
          axis.text.x = element_text(size = 7),
          axis.text.y = element_text(size = 7),
          axis.title.x = element_text(size = 7),
          axis.title.y = element_text(size = 7),
          plot.margin = margin(0, 0, 0, 0))

}

centered_data <- center_data(cone)

## First projection
projection <- cbind(
  c(0.22417,-0.02943,-0.02870,0.08418),
  c(-0.02715,0.18019,0.05053,0.15254))

proj_obj1 <- get_projection(projection = projection, 
                            centered_data = centered_data, 
                            axis_param = list(limits = 0.6,
                                              axis_scaled = 3, 
                                              axis_pos_x = -0.4, 
                                              axis_pos_y = -0.4, 
                                              threshold = 0.053))

cone_proj1 <- plot_proj(
  proj_obj = proj_obj1, 
  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
  plot_limits = c(-0.5, 0.7), 
  title = "a1", 
  cex = 2, 
  axis_text_size = 5,
  is_color = FALSE)

## Second projection
projection <- cbind(
    c(0.15591,0.17907,0.03198,0.04496),
    c(-0.15376,0.10435,0.15766,0.00547))

proj_obj2 <- get_projection(projection = projection, 
                            centered_data = centered_data, 
                            axis_param = list(limits = 0.3,
                                              axis_scaled = 2, 
                                              axis_pos_x = -0.2, 
                                              axis_pos_y = -0.2, 
                                              threshold = 0.02))

cone_proj2 <- plot_proj(
  proj_obj = proj_obj2, 
  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
  plot_limits = c(-0.25, 0.35), 
  title = "a2", 
  cex = 2, 
  axis_text_size = 5,
  is_color = FALSE)

## Third projection
projection <- cbind(
    c(-0.13548,0.13832,0.05991,-0.13544),
    c(0.08479,-0.10719,-0.01572,-0.20123))

proj_obj3 <- get_projection(projection = projection, 
                            centered_data = centered_data, 
                            axis_param = list(limits = 0.6,
                                              axis_scaled = 3, 
                                              axis_pos_x = -0.6, 
                                              axis_pos_y = -0.6, 
                                              threshold = 0.04))

cone_proj3 <- plot_proj(
  proj_obj = proj_obj3, 
  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
  plot_limits = c(-0.7, 0.5), 
  title = "a3", 
  cex = 2, 
  axis_text_size = 5,
  is_color = FALSE)

cone_proj1 + cone_proj2 + cone_proj3 +
  plot_layout(ncol = 3, guides = "collect") 
```