### 2025-12-19 extra/paper-cardinalR-code.Rmd
Updated with new package versions

---
title: "Sample cardinalR"
author: "Chat-GPT"
date: "2025-09-10"
format: pdf
---

### Branching

A branching structure (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch)"} else { "Figure \\@ref(fig:branch-proj)"})`) captures trajectories that diverge or bifurcate from a common origin, similar to processes such as cell differentiation in biology [@trapnell2014]. We introduce a set of data generation functions specifically designed to simulate high-dimensional branching structures with various geometries, total number of points (`n`) generated across all branches, with points allocated approximately evenly among branches, and number of branches (`k`). Although these functions can generate multiple branches, they do not produce a formal *multicluster* dataset: the branches form a single connected structure, with multiple visually distinct arms rather than independent clusters.

The simplest structures are approximately linear branches in $2\text{-}D$, generated by the `gen_linearbranches(n, k)` function. These consist of $k$ short line segments in the first two dimensions, with added jitter to simulate variability. Mathematically, each branch $i$ is defined as

$$
X_1 \sim U(a_i, b_i), \quad X_2 = s_i (X_1 - x_{\text{start},i}) + y_{\text{start},i} + \epsilon, \quad \epsilon \sim U(0, \delta),
$$

where $(x_{\text{start},i}, y_{\text{start},i})$ is the starting point of branch $i$, $\delta$ controls local jitter, and $s_i$ is the slope, initialized as

$$
s_i =
\begin{cases}
0.5 & i = 1, \\
-0.5 & i = 2, \\
\text{randomly sampled from } [s_{\min}, s_{\max}] & i = 3, \dots, k.
\end{cases}
$$

The jitter term is sampled from a one-sided uniform distribution to introduce directional variability without altering branch orientation.

Branches $1$ and $2$ are initialized with fixed slopes and starting points, while later branches are iteratively added at locations chosen to avoid overlap with existing branches, producing a set of connected linear paths.

To introduce curvature, the `gen_curvybranches(n, k)` function generates $k$ curvilinear branches in $2\text{-}D$. Each branch follows a quadratic trajectory of the form

$$
X_1 \sim U(a_i, b_i), \quad X_2 = 0.1 X_1 + s_i X_1^2 + \epsilon, \quad \epsilon \sim U(-\delta, \delta),
$$
where $(a_i, b_i)$ defines the domain of the branch, $s_i$ controls curvature, and $\delta$ introduces local jitter. For the first two branches, the parameters are fixed to establish reference shapes: $(a_1,b_1,s_1) = (0,1,1), \quad (a_2,b_2,s_2) = (-1,0,-2)$. Additional branches are attached iteratively to existing structures. Each new branch $i$ starts at a selected point $(x_{\text{start},i}, y_{\text{start},i})$ from the current structure and extends according to

$$
X_1 \sim U(x_{\text{start},i}, x_{\text{start},i}+1), \quad X_2 = 0.1 X_1 - s_i (X_1^2 - x_{\text{start},i}) + y_{\text{start},i},
$$

where $s_i$ is a scale factor controlling the curvature of branch $i$. For the first few initial branches, $s_i$ can be fixed (e.g., $s_1 = 1, s_2 = 2$) to establish reference shapes, while for subsequent branches it is sampled from a predefined set, such as $s_i \in \{-2, -1.5, -1, -0.5, 0, 0.5, 1, 1.5\}$, to control curvature magnitude and direction.

The `gen_expbranches(n, k)` function creates $k$ exponential branches in $2\text{-}D$, radiating from a central region. Each branch $i$ is defined as

$$
X_1 \sim U(-2,2), \quad X_2 = \exp(\sigma_i \, s_i \, X_1) + \epsilon, \quad \epsilon \sim U(0, \delta), \quad s_i \sim U(0.5,2),
$$

where $\sigma_i = (-1)^{i+1}$ alternates the sign of the exponent to produce mirror-symmetric branches. The parameter $s_i$ controls the steepness of branch $i$, and $\delta$ introduces small local jitter.

High-dimensional generalizations are provided by `gen_orglinearbranches(n, p, k)` (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:branch)"} else { "Figure \\@ref(fig:branch-proj)"})`) and `gen_orgcurvybranches(n, p, k)`. For branch 
$i$, the active coordinate pair $(i_1, i_2)$ indexes the selected $2\text{-}D$ subspace. When `allow_share = TRUE`, multiple branches may share the same subspace; otherwise, subspaces are sampled without replacement until all possible $\binom{p}{2}$ combinations are exhausted, after which additional branches may repeat subspaces.

In both cases, branch $i$ is generated according to

$$
X_{i_1} \sim U(a_i, b_i), \quad
X_{i_2} = f_i(X_{i_1}) + \epsilon, \quad
\epsilon \sim N(0, \sigma^2),
$$
where $a_i$ and $b_i$ define the domain of the branch and $\epsilon$ introduces smooth variability in the $p\text{-}D$ space. The function $f_i(\cdot)$ determines the branch geometry:
$$
f_i(x) =
\begin{cases}
s_i x, & \text{linear branches}, \\
-s_i x^2, & \text{curvilinear branches}.
\end{cases}
$$

The scale factor $s_i$ controls slope (linear branches) or curvature (curvilinear branches) and is assigned as follows: for the first $\binom{p}{2}$ branches, $s_i = 1$; for additional branches when $k > \binom{p}{2}$, $s_i$ is randomly drawn from the set $\{1, 1.5, 2, \dots, 8\}$.

Across all branching generators, the scale parameter $s_i$ controls the strength of deviation from linearity, determining slope, curvature, or growth rate depending on the branch geometry.

```{r branch-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from the $4\\text{-}D$ \\texttt{orgcurvybranches} data. Each shows a different projection, illustrating how the linear branches appear from multiple viewing angles. These views highlight the dataset’s underlying branching structure and demonstrate how projections reveal patterns that are otherwise hidden in higher dimensions.", fig.width=12, fig.height=4, fig.pos="H", fig.pos="!ht"}

set.seed(20240412)

# Add plot label

interior_annotation <- function(label, position = c(0.92, 0.92), cex = 1, col="grey70") {
  annotation_custom(grid::textGrob(label = label,
                                   x = unit(position[1], "npc"), y = unit(position[2], "npc"),
                                   gp = grid::gpar(cex = cex, col=col)))
}

# Scale high-d data

# Center the data by subtracting the mean of each column
center_data <- function(data) {
  center_values <- colMeans(data)
  data_centered <- sweep(data, 2, center_values, FUN = "-")  # subtract means
  data_centered
}

# Get projection

get_projection <- function(projection, centered_data, axis_param) {

  projected <- as.matrix(centered_data) %*% projection
  projected_df <- projected |>
    tibble::as_tibble(.name_repair = "unique") |>
    dplyr::rename(c("proj1" = "...1",
                    "proj2" = "...2")) |>
    dplyr::mutate(ID = dplyr::row_number())

  limits <- axis_param$limits
  axis_scaled <- axis_param$axis_scaled
  axis_pos_x <- axis_param$axis_pos_x
  axis_pos_y <- axis_param$axis_pos_y
  threshold <- axis_param$threshold

  axes_obj <- gen_axes(
    proj = projection * axis_scaled,
    limits = limits,
    axis_pos_x = axis_pos_x,
    axis_pos_y = axis_pos_y,
    axis_labels = names(centered_data),
    threshold = threshold)

  axes <- axes_obj$axes
  circle <- axes_obj$circle

  return(list(projected_df = projected_df,
              axes = axes,
              circle = circle))

}

# Plot projection
plot_proj <- function(proj_obj,
                      point_param = c(1.5, 0.5, "#000000"), # size, alpha, color
                      plot_limits, title, cex = 2,
                      position = c(0.92, 0.92),
                      axis_text_size = 3,
                      is_color = FALSE) {

  projected_df <- proj_obj$projected_df
  axes <- proj_obj$axes
  circle <- proj_obj$circle

  if(is_color == FALSE) {

    initial_plot <- ggplot() +
      geom_point(
        data = projected_df,
        aes(
          x = proj1,
          y = proj2),
        size = as.numeric(point_param[1]),
        alpha = as.numeric(point_param[2]),
        color = point_param[3])

  } else {

    projected_df <- projected_df |>
      dplyr::mutate(cluster = proj_obj$cluster)

    initial_plot <- ggplot() +
      geom_point(
        data = projected_df,
        aes(
          x = proj1,
          y = proj2,
          color = cluster),
        size = as.numeric(point_param[1]),
        alpha = as.numeric(point_param[2]))

  }

  initial_plot <- initial_plot +
    geom_segment(
      data=axes,
      aes(x=x1, y=y1, xend=x2, yend=y2),
      colour="grey70") +
    geom_text(
      data=axes,
      aes(x=x2, y=y2),
      label=rownames(axes),
      colour="grey50",
      size = axis_text_size) +
    geom_path(
      data=circle,
      aes(x=c1, y=c2), colour="grey70") +
    xlim(plot_limits) +
    ylim(plot_limits) +
    interior_annotation(title, position, cex = cex)

  initial_plot

}

# Generate axes

gen_axes <- function(proj, limits = 1, axis_pos_x = NULL, axis_pos_y = NULL, axis_labels, threshold) {

  axis_scale <- limits/6

  if (is.null(axis_pos_x)) {

    axis_pos_x <- -2/3 * limits

  }

  if (is.null(axis_pos_y)) {

    axis_pos_y <- -2/3 * limits

  }

  adj <- function(x, axis_pos) axis_pos + x * axis_scale
  axes <- data.frame(x1 = adj(0, axis_pos_x),
                     y1 = adj(0, axis_pos_y),
                     x2 = adj(proj[, 1], axis_pos_x),
                     y2 = adj(proj[, 2], axis_pos_y))

  rownames(axes) <- axis_labels

  ## To remove axes
  axes <- axes |>
    mutate(distance = sqrt((x2 - x1)^2 + (y2 - y1)^2)) |>
    filter(distance >= threshold)

  theta <- seq(0, 2 * pi, length = 50)
  circle <- data.frame(c1 = adj(cos(theta), axis_pos_x),
                       c2 = adj(sin(theta), axis_pos_y))

  return(list(axes = axes, circle = circle))

}

plot_hbe <- function(error_df) {

  ggplot(error_df,
         aes(x = a1,
             y = HBE,
             colour = method)) +
    geom_point(size = 0.8) +
    geom_line(linewidth = 0.3) +
    ylab("HBE") +
    xlab(expression(paste("binwidth (", a[1], ")"))) +
    theme_minimal() +
    theme(panel.border = element_rect(fill = 'transparent'),
          plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
          axis.ticks.x = element_line(),
          axis.ticks.y = element_line(),
          legend.position = "none",
          axis.text.x = element_text(size = 7),
          axis.text.y = element_text(size = 7),
          axis.title.x = element_text(size = 7),
          axis.title.y = element_text(size = 7),
          plot.margin = margin(0, 0, 0, 0))

}

orglinearbranches <- gen_orglinearbranches(n = 1000, p = 4, k = 4)

## First projection
projection <- cbind(
    c(-0.083573,-0.019043,0.006574,-0.035774),
    c(0.005380,0.068541,-0.030810,-0.054717))

proj_obj1 <- get_projection(projection = projection, 
                            centered_data = centered_data, 
                            axis_param = list(limits = 0.5,
                                              axis_scaled = 10, 
                                              axis_pos_x = -0.5, 
                                              axis_pos_y = -0.5, 
                                              threshold = 0.05))

orglinearbranches_proj1 <- plot_proj(
  proj_obj = proj_obj1, 
  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
  plot_limits = c(-0.6, 0.4), 
  title = "a1", 
  cex = 2, 
  axis_text_size = 4,
  is_color = FALSE)

## Second projection
projection <- cbind(
  c(-0.06614,0.15134,-0.44278,0.27367),
  c(-0.50493,0.14388,0.14615,0.03486))

proj_obj2 <- get_projection(projection = projection, 
                            centered_data = centered_data, 
                            axis_param = list(limits = 0.9,
                                              axis_scaled = 1, 
                                              axis_pos_x = -0.85, 
                                              axis_pos_y = -0.85, 
                                              threshold = 0.05))

orglinearbranches_proj2 <- plot_proj(
  proj_obj = proj_obj2, 
  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
  plot_limits = c(-1, 1), 
  title = "a2", 
  cex = 2, 
  axis_text_size = 4,
  is_color = FALSE)

## Third projection
projection <- cbind(
  c(0.10704,0.04822,-0.03416,0.53224),
  c(0.50329,0.07199,-0.16064,-0.11806))

proj_obj3 <- get_projection(projection = projection, 
                            centered_data = centered_data, 
                            axis_param = list(limits = 0.9,
                                              axis_scaled = 1.5, 
                                              axis_pos_x = -0.85, 
                                              axis_pos_y = -0.85, 
                                              threshold = 0.1))

orglinearbranches_proj3 <- plot_proj(
  proj_obj = proj_obj3, 
  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
  plot_limits = c(-1, 1), 
  title = "a3", 
  cex = 2, 
  axis_text_size = 4,
  is_color = FALSE)

orglinearbranches_proj1 + orglinearbranches_proj2 + orglinearbranches_proj3 +
  plot_layout(ncol = 3, guides = "collect") 
```